### 2023-03-02 paper/admin/cover_letter.tex
started to work on cover letter (backing up progress since i need to step away)

\title{cover letter}
%
% See http://texblog.org/2013/11/11/latexs-alternative-letter-class-newlfm/
% and http://www.ctan.org/tex-archive/macros/latex/contrib/newlfm
% for more information.
%
\documentclass[11pt,stdletter,orderfromtodate,sigleft]{newlfm}
\usepackage{blindtext, xfrac, animate, hyperref, pxfonts, geometry}

  \setlength{\voffset}{0in}

\newlfmP{dateskipbefore=0pt}
\newlfmP{sigsize=20pt}
\newlfmP{sigskipbefore=10pt}

\newlfmP{Headlinewd=0pt,Footlinewd=0pt}

\newcommand{\journal}{Nature Neuroscience}
\newcommand{\articletype}{Article}
\newcommand{\myTitle}{High-order cognition is supported by information-rich but compressible brain activity patterns}
\newcommand{\corresponding}{Jeremy R. Manning}

\namefrom{\vspace{-0.3in}\corresponding}
\addrfrom{
  Dartmouth College\\
  Department of Psychological \& Brain Sciences\\
  Moore Hall\\
  Hanover, NH, 03755}
\addrto{}
\dateset{\today}

\greetto{To the editors of \textit{\journal}:}

\closeline{Sincerely,}
%\usepackage{setspace}
%\linespread{0.85}
% How will your work make others in the field think differently and move the field forward?
% How does your work relate to the current literature on the topic?
% Who do you consider to be the most relevant audience for this work?
% Have you made clear in the letter what the work has and has not achieved?

\begin{document}
\begin{newlfm}

  We have enclosed our manuscript entitled \textit{\myTitle} to be considered
  for publication as an \textit{\articletype}. Our manuscript uses
  dimensionality reduction algorithms and pattern classifiers to explore how
  informative and compressible (in the information theory senses of those
  words) brain activity patterns are, under different cognitive circumstances.

  We developed a computational framework for evaluating our two measures of
  interest (informativeness and compressibility), and we applied the framework
  to a public fMRI dataset. In the dataset's experiment, participants either
  listened to an auditory recording of a story, listened to temporally
  scrambled recordings of the story, or underwent a resting state scan. Each
  condition is designed to engage cognitive processing and engagement at
  different depths. For example, listening to the intact story leads
  participants to mentally engage with the narrative, and to deeply process
  and connect narrative events in a way that leads to a rich understanding
  of the story.  Listening to temporally scrambled versions of the story might
  lead to understanding invidual words....

Thank you for considering this manuscript, and I hope you will find it suitable
for publication in \textit{\journal}.

\end{newlfm}
\end{document}