### 2023-03-15 paper/figs/source/neurosynth_topics.tex
starting to add material to supplement, also added new figure to main text and removed the old neurosynth figures

\begin{tabular}{rlllllllllll}
\toprule
                                       Topic Label &     Cognitive Label &           Term 1 &        Term 2 &          Term 3 &         Term 4 &      Term 5 &         Term 6 &        Term 7 &         Term 8 &        Term 9 &        Term 10 \\
\midrule
            Cognitive control and task performance &   Cognitive control &            tasks &       control &         network &     conditions &  comparison &      performed &        common &     correlates &    experiment &            pre \\
                Developmental aging and maturation &                   - &              age &        adults &        children &    development & adolescents &          aging & developmental &    adolescence &     childhood &          adult \\
                Eye movements and visual attention &           Attention &              eye &          gaze &            eyes &         visual &    saccades &      movements &       saccade &      direction &          gait &         target \\
                      Facial and voice recognition &  Sensory perception &      recognition &      familiar &        identity &     unfamiliar &       voice &    familiarity &         route &         voices &            fg &         facial \\
        Social interaction and contextual behavior &    Social cognition &          context &          game &           human &    interaction &         ppi &     contextual &      contexts &         agency &  interactions &        partner \\
        Language processing and semantic knowledge & Language processing &         semantic &         words &            word &        lexical &      verbal &       language &         tasks &         naming &       fluency &   phonological \\
    Experimental design and behavioral performance &                   - &           trials &      stimulus &       responses &          trial &    reaction &           time &         event &         target &        events &          times \\
            Genetic polymorphisms and risk factors &                   - &         carriers &        allele &            gene &       genotype &         met &        genetic &  polymorphism &            val &          comt &             rs \\
     Sensorimotor integration and movement control &       Motor control &            motor &      movement &       movements &   sensorimotor &     primary &         finger &       control &        imagery &       sensory &      execution \\
                Drug addiction and substance abuse &                   - &          cocaine &         users &            drug &            bpd &    controls &       cannabis &     addiction &        craving &     dependent &         heroin \\
          Music perception and auditory processing &  Sensory perception &            music &       musical &           pitch &       auditory &   musicians &      sequences &        rhythm &      listening &          beat &        singing \\
           Menstrual cycle and hormonal regulation &                   - &            phase &         women &           cycle &         phases &   menstrual &             hf &    expression &            sex &        luteal &     follicular \\
              Cognitive functions and role playing &   Cognitive control &             role &          play &           human &          plays &   cognitive &       evidence &      critical &       distinct &           key &     subregions \\
                 Inhibition and gender differences &                   - &       inhibition &         women &      inhibitory &            sex &     females &         gender &         males &           stop &          male &         female \\
       Somatosensory stimulation and motor control &       Motor control &      stimulation & somatosensory &             tms &        tactile &     primary &          touch &         motor &           rtms &  transcranial &        sensory \\
           Multisensory integration and perception &  Sensory perception &         auditory &        visual &         sensory &       modality &      sounds &    integration &         sound &       stimulus &       primary &     modalities \\
                     Social perception and empathy &    Social cognition &           social &       empathy &      experience &         people &      person &      responses &   perspective &    individuals &    attachment &       empathic \\
          Gesture recognition and visual attention &           Attention &           target &      gestures &         targets &    orientation &      visual &    distractors &       gesture &          grasp &    distractor &       location \\
                               Experimental design &                   - &           design &         block &          blocks &          event &       mixed &      condition &            ca &        writing &       blocked &           runs \\
                            Alcohol cue reactivity &              Reward &             cues &           cue &         alcohol &   anticipation &        cued &    preparation &  anticipatory &       exposure &    expectancy &    preparatory \\
                       Neuroimaging and metabolism &                   - &              pet &    tomography &        emission &       positron &        flow &        glucose &       binding &     metabolism &     metabolic &       receptor \\
                    Abnormalities in schizophrenia &                   - &    schizophrenia &      controls &         reduced &  abnormalities &    symptoms &       deficits &       matched &        control &      abnormal &  schizophrenic \\
                            Eating and body weight &                   - &             food &         taste &            body &         weight &      eating &          women &         obese &         reward &         foods &        caloric \\
                    Sleep and olfactory processing &  Sensory perception &            sleep &     olfactory &            odor &             sd & deprivation &          odors &           rem &    wakefulness &         night &           wake \\
 Alzheimer's disease and mild cognitive impairment &                   - &               ad &       disease &             mci &      alzheimer &   cognitive &     impairment &          mild &           amci &      controls &        atrophy \\
             Working memory and executive function &              Memory &           memory &          load &           tasks &         verbal & maintenance &    performance &          term &     difficulty &          dual &      cognitive \\
                 Moral decision making and phobias &                   - &            moral &        phobia &          phobic &             ec &       guilt &         spider &      decision &            art &       phobics &    intentional \\
                               Language laterality & Language processing &         language &     asymmetry &    organization &          human &   dominance &    asymmetries &   lateralized &         handed &       located & representation \\
                                         Attention &           Attention &        attention &   attentional &          visual &        spatial &         top &      selective &      stimulus &        control &     orienting &         target \\
           Resting-state brain activity in smokers &       Resting state &             reho &       smokers &         resting &        smoking &    nicotine &    homogeneity &           sci &       controls &            rs &    spontaneous \\
                         Social cognition/judgment &    Social cognition &           social &     judgments &            mind &         mental &      theory &    mentalizing &     cognition &       judgment &        person &         people \\
                        Reward and decision making &              Reward &           reward &      decision &          choice &       outcomes &     rewards &       monetary &     decisions &   anticipation &     responses &        outcome \\
                       ADHD and attention deficits &           Attention &             adhd &      disorder &       attention &        deficit &    children &  hyperactivity &      deficits &        control &      controls &        reduced \\
Neurobiological variability and individual diff... &                   - &       individual &  relationship &           local &      dependent &      change &         global &     responses &       neuronal &     magnitude &          lower \\
                                 Spatial cognition &   Spatial cognition &          spatial &         space &        location &      locations &  navigation &        virtual &        visual &   visuospatial &         visuo &       position \\
            Therapeutic interventions and training &                   - &         training &   acupuncture &         therapy &             cr &     control &        trained &   improvement &    mindfulness &   stimulation &        induced \\
                    Color perception and deception &  Sensory perception &            color &        search &         feature &      deception &    features &      responses &        colour &      dimension &         lying &    conjunction \\
          Neurodegenerative diseases and disorders &                   - &          disease &            pd &        controls &        atrophy &    clinical &          motor &      dementia &             sd &      multiple &      sclerosis \\
                Cognitive control and interference &   Cognitive control &         conflict &       control &    interference &         stroop & incongruent &      selection &     congruent &         trials &     cognitive &     monitoring \\
          Structural MRI and brain volume analysis &                   - &           volume &          gray &           voxel &             gm & morphometry &           grey &       volumes &            vbm &       density &            age \\
                  Fear conditioning and extinction &             Emotion &             fear &     switching &    conditioning &      responses &    stimulus &     extinction &            cs &         switch &        threat &    conditioned \\
                      Skill learning and expertise &              Memory &         learning &      practice &         learned &       sequence & performance &       training &     sequences &          skill &      implicit &          motor \\
                                   PTSD and trauma &             Emotion &             ptsd &        trauma &          stress &       disorder &   traumatic &  posttraumatic &     childhood &      survivors &      exposure &       controls \\
         Neural oscillations and electrophysiology &                   - &        frequency &        source &           alpha &      amplitude &        beta &          gamma &      recorded &    frequencies &     potential &   simultaneous \\
          Temporal dynamics of stimulus processing &  Sensory perception &             time &     sustained &        duration &          onset &      period &          stage &        timing &          delay &     transient &          event \\
                         Tinnitus and hearing loss &  Sensory perception &         tinnitus &          loss &         hearing &         status &     driving &     subjective &     objective &         unfair &        offers &      rejection \\
           Abstract categories and representations & Language processing &         category &    adaptation & representations & categorization &  categories &       abstract &      stimulus & representation &      features &      knowledge \\
           Pain perception and sensory stimulation &  Sensory perception &             pain &       painful &     stimulation &  somatosensory &   intensity &        noxious &          heat &        chronic &       sensory &    nociceptive \\
                                 Body and primates &                   - &             body &         human &          humans &        monkeys &        itch &       primates &       species &         monkey &        bodies &        macaque \\
                Phonological processing in reading & Language processing &          reading &       chinese &    phonological &         visual &    language &        readers &      dyslexia &     characters &      children &           word \\
             Rule-based performance and complexity &   Cognitive control &             rule &    complexity &            size &          force &       rules &         effect &          grip &         amount &      distance &     artificial \\
Autism Spectrum Disorder (ASD) and social impai... &    Social cognition &              asd &        autism &          social &       spectrum & individuals &       controls &     disorders &       children &            td &        reduced \\
            Major depression disorder and emotions &             Emotion &       depression &           mdd &       depressed &     depressive &       major &           mood &      disorder &            sad &      controls &        control \\
                              Blindness and vision &  Sensory perception &            blind &        visual &         sighted &    individuals &       humor &       laughter &  congenitally &      blindness &    plasticity &        braille \\
                        Deafness and sign language & Language processing &        condition &    conditions &            deaf &        hearing &        sign &          signs &   referential &        signers &            ci &            nss \\
    Genetic risk and familial factors in psychosis &                   - &             risk &       genetic &              sz &       siblings &   relatives &    individuals &    unaffected &        factors &        family &      psychosis \\
                  Action observation and imitation &       Motor control &           action &       actions &     observation &          motor &      mirror &           goal &     imitation &      execution &      directed &      movements \\
                 Cognitive performance and control &   Cognitive control &      performance &     cognitive &         control &         memory &   executive &           test &         tasks &    individuals &    behavioral &       deficits \\
                     Mental disorders and controls &                   - &         disorder &           ocd &         bipolar &             hc &          bd &       controls &            ts &        control & abnormalities &     compulsive \\
Pharmacological effects of placebo and drug adm... &                   - &          placebo &      dopamine &          effect &           drug &          ht & administration &         blind &   dopaminergic &        double &             mg \\
                           Personality and anxiety &    Social cognition &          anxiety &   personality &           trait &    individuals &      scores &         traits &        threat &      disorders &        social &      avoidance \\
                 Mental imagery and math abilities &   Cognitive control &           mental &       imagery &       numerical &     arithmetic &    rotation &    calculation &     magnitude &          digit &         tasks &   mathematical \\
                     Priming and repetition effect &              Memory &          priming &    repetition &     suppression &       repeated &      effect &       implicit &       literal &         target &         prime &      metaphors \\
               Working memory and error monitoring &              Memory &               wm &         error &          errors &     prediction & performance &        correct &    monitoring &            ltm &      feedback &      predicted \\
                 Sentence comprehension and syntax & Language processing &        sentences &      language &        sentence &  comprehension &   syntactic &       semantic &         verbs &           verb &          word &        meaning \\
                            Resting state networks &       Resting state &          network &       resting &         default &           mode &        rest &      intrinsic &     cognitive &   correlations &          seed &    spontaneous \\
            Episodic memory encoding and retrieval &              Memory &           memory &      encoding &       retrieval &    recognition &    episodic &          items &    successful &     subsequent &  recollection &         recall \\
                         Visual object recognition &  Sensory perception &           object &        visual &         objects &          shape &      images &         scenes &         scene &      selective &   recognition &         stream \\
      Effective causal modeling of neural networks &                   - &        effective &        causal &        modeling &        dynamic &     network &            top &     influence &     modulation &  interactions &      causality \\
       Relational reasoning and fluid intelligence &                   - &        reasoning &         FALSE &            TRUE &   intelligence &  relational &         belief &     relations &      nonverbal &         fluid &     analogical \\
                  Lesion and stroke rehabilitation &                   - &          lesions &        lesion &          stroke &         damage &      injury &       controls &      recovery &        patient &        tracts &      integrity \\
         Affective valence and feedback processing &                   - &         negative &      positive &        feedback &        valence &     arousal &           bias &    positively &     negatively &    evaluation &     swallowing \\
               Autobiographical memory in epilepsy &              Memory & autobiographical &        events &          future &       epilepsy &    episodic &         memory &      memories &           past &          mtle &       personal \\
         Evidence and effect in behavioral studies &                   - &         evidence &       provide &          effect &     behavioral &  underlying &   demonstrated & understanding &      potential &     responses &        support \\
                Stress and physiological responses &                   - &           stress &          tdcs &        cortisol &      autonomic &       heart &      responses &          rate &     regulation & physiological &        induced \\
                    Speech and language processing & Language processing &           speech &      language &        auditory &     production &  perception &  comprehension &     listening &       acoustic &    linguistic &        prosody \\
Network interactions and evidence in human systems &                   - &          network &      evidence &           human &        systems &     support &        process &      distinct &    integration &       provide &        engaged \\
                           Neuroimaging techniques &                   - &         standard &        images &      individual &           time &       image &          voxel &       spatial &           test &      clinical &        mapping \\
              Visual perception of motion and form &  Sensory perception &           motion &        visual &      perception &     perceptual &  biological &        dynamic &        moving &          human &        static &       illusion \\
               Emotional processing and regulation &             Emotion &        emotional &       emotion &         neutral &         facial & expressions &      affective &     responses &       negative &    regulation &       emotions \\
\bottomrule
\end{tabular}